Predicting Spotify Audio Features from Last.fm Tags

### 2023-06-19 paper/main.tex
apply suggestions from Philippe

Similarly to Last.fm, Spotify (or Echo Nest) audio features are commonly present in MIR research.
For example, Wang and Horv{\'a}t used  these audio features to study the differences 
between male and female artists \cite{wang2019gender}.
\textcolor{red}{Y que consiguieron?}

@@

For this reason, we chose to generate our own dataset.

@@

% Additionally, language models, such as GTP-2 have shown promising results in various
% natural language processing (NLP) and generation.

% In this particular study, GTP-2 has been fine-tuned for regression tasks,
% and has shown good performance in predicting audio features from metadata.

% This paper aims to apply the boosted tree regressor,
%  Bayesian ridge regressor, and GTP-2 models to predict Spotify audio features from Last.fm tags.
% The experiments compare the performance of these models and evaluate their effectiveness
% in predicting various audio features.
% The results of these experiments will provide insights into
% the use of different machine learning models for MIR tasks
% and can have practical applications
% in music recommendation systems and genre or mood recognition.
%
%\textcolor{red}{To be Continued \dots }

@@

Users tipically listen to their favorite tracks several times,
so the amount of unique tracks played is smaller
than the number of track plays. In this case, the amount of
individual tracks listened in a 15-year period is about \num{20000} and the number of scrobblings is, approximately, \num{90000}.
Therefore, the user has listened to each song, approximately, \num{4.5} times on average.

@@

\end{table}
% \begin{table}[h!]
%       \begin{center}

%       \end{center}
% \end{table}

@@

@@

The \emph{GTP-2} model is a transformer model.
Transformers are a type of neural network architecture that has been widely used in natural language processing (NLP) tasks,
such as text generation and question answering.
GPT-2 has been trained on a large corpus of text, and typically used as a language model to generate text that resembles human-written language.
However, in this study we have experimented with GPT-2 as a regressor.
Rather than retraining the model from scratch, we have fine tuned the model to behave as a regressor and  predict Spotify audio features.
The basic idea behind this approach is to feed a string of concatenated Last.fm tags as input to the GPT-2 model, and then use the model's output as the predicted value of a specific Spotify audio feature.

@@

The preceding models require specific input formats.
In particular, the boosted tree and bayesian ridge regressors require structured data (e.g a set of predictor features and a set of target variables), whereas GTP-2 expects text strings as input.

@@

% Tags are converted to text tokens. Columns represent token positions, and cells contain the token at a particular position, for a track.
% To tokenize tags, we have used the GTP-2 tokenizer.

@@

Similarly to the tags tabular format, we also defined fixed values for the number of columns: 10, 1,000, and 10,000.
\textcolor{red}{Decir cual es el nombre de este limite (maxsequence length?)}

@@

% We configured the boosted tree regressor model with the training parameters listed in table \ref{table_xgboost_training_params}.

% \begin{table}[h!]
%       \begin{center}
%       \begin{minipage}{174pt}
%       \caption{Training parameters for XGBoost regressor}\label{table_xgboost_training_params}%
%       \begin{tabular}{@{}llll@{}}
%       \toprule
%       Parameter               & Value \\
%       \midrule
%       objective               & reg:squarederror  \\
%       base score              & 0.5 \\
%       booster                 & gbtree  \\
%       colsample bylevel       & 1 \\
%       colsample bynode        & 1 \\
%       colsample bytree        & 1 \\
%       gamma\footnotemark[1]   & 0 \\
%       learning rate           & 0.300000012 \\
%       max delta step          & 0 \\
%       max depth               & 6 \\
%       min child weight        & 1 \\
%       estimators              & 200  \\
%       n jobs                  & 12  \\
%       num parallel tree       & 1 \\
%       predictor               & auto  \\
%       random state            & 0 \\
%       reg alpha               & 0 \\
%       reg lambda              & 1 \\
%       scale pos weight        & 1 \\
%       subsample               & 2 \\
%       tree method             & auto  \\
%       \botrule
%       \end{tabular}
%       \footnotetext[1]{Minimum loss reduction required to make a further partition on a leaf node of the tree.}
%       \end{minipage}
%       \end{center}
% \end{table}

@@

% The training parameters are listed in table \ref{table_bayesian_training_params}.

% \begin{table}[h]
%       \begin{center}
%       \begin{minipage}{174pt}
%       \caption{Training parameters for XGBoost regressor}\label{table_bayesian_training_params}%
%       \begin{tabular}{@{}llll@{}}
%       \toprule
%       Parameter                     & Value \\
%       \midrule
%       Maximum iterations            & 300   \\
%       Tolerance\footnotemark[1]     & \num{1e-03} \\
%       alpha 1                       & \num{1e-06} \\
%       alpha 2                       & \num{1e-06} \\
%       lambda 1                      & \num{1e-06} \\
%       lambda 2                      & \num{1e-06} \\
%       \botrule
%       \end{tabular}
%       \footnotetext[1]{Tolerance for the stopping criteria.}
%       \end{minipage}
%       \end{center}
% \end{table}

@@

The experiments tested the three models, and the three values of $K$: \num{100}, \num{1000}, \num{10000}.
In the Bayesian Ridge and Boosted tree regressors, we tested the two structured data formats: tags and tokens as columns.
In the GTP-2 model, we tested the text data format.

@@

@@

The GTP-2 model was the best model to predict the danceability, with a \num{0.145} RMSE.
The Boosted tree was the best model to predict the instrumentalness, with a RMSE value of \num{0.29}.

@@

The Duplicated stringification method consistently performed worse with the GTP-2 transformer model.

>>>>

Similarly to Last.fm, Spotify (or Echo Nest) audio features are commonly present in MIR research.
For example, Wang and Horv{\'a}t used  these audio features to study the differences 
between male and female artists \cite{wang2019gender}.
\textcolor{red}{And what did they achieve?}

@@

There are various reasons why we chose to generate our own dataset.
First, we wanted to generate a dataset that combined both Last.fm tags and Spotify audio features.
Second, there is lack of clarity regarding the conditions under which Spotify allows the use of the audio features.
Third, we wanted to explore a research line where machine learning models are trained by using the listening history of a single user. 
% Another dataset created from the Spotify API and published to Kaggle:
% https://www.kaggle.com/datasets/andrewmvd/spotify-playlists

@@

% Additionally, language models, such as GPT-2 have shown promising results in various
% natural language processing (NLP) and generation.

% In this particular study, GPT-2 has been fine-tuned for regression tasks,
% and has shown good performance in predicting audio features from metadata.

% This paper aims to apply the boosted tree regressor,
%  Bayesian ridge regressor, and GPT-2 models to predict Spotify audio features from Last.fm tags.
% The experiments compare the performance of these models and evaluate their effectiveness
% in predicting various audio features.
% The results of these experiments will provide insights into
% the use of different machine learning models for MIR tasks
% and can have practical applications
% in music recommendation systems and genre or mood recognition.
%
%\textcolor{red}{To be Continued \dots }

@@

Users typically listen to their favorite tracks several times,
so the amount of unique tracks played is smaller
than the number of track plays. In this case, the amount of
individual tracks listened in a 15-year period is about \num{20000} and the number of scrobblings is, approximately, \num{90000}.
Therefore, the user has listened to each song, approximately, \num{4.5} times on average.

@@

\end{table}

@@

\subsection{Models}

@@

The \emph{GPT-2} model is a transformer model.
Transformers are a type of neural network architecture that has been widely used in natural language processing (NLP) tasks,
such as text generation and question answering.
GPT-2 has been trained on a large corpus of text, and typically used as a language model to generate text that resembles human-written language.
However, in this study we have experimented with GPT-2 as a regressor.
Rather than retraining the model from scratch, we have fine tuned the model to behave as a regressor and  predict Spotify audio features.
The basic idea behind this approach is to feed a string of concatenated Last.fm tags as input to the GPT-2 model, and then use the model's output as the predicted value of a specific Spotify audio feature.

\subsubsection{Models Parametrization}

Due to the different models and input encodings that we employed in the models, we decided to limit the dimensionality of the experiments, 
by using mostly the default hyperparameters that the most common libraries set for these models.

For the Bayesian ridge regressor, we used the default hyperparameters set by the \emph{scikit-learn} library.
The training parameters for the Bayesian ridge are listed in table \ref{table_bayesian_training_params}.

To fine tune the GPT-2 model, we used the \emph{GPT2ForSequenceClassification} model of the \emph{Transformers} Python library.
The parameters used are listed in table \ref{table_gpt2_training_params}.

\begin{table}[!ht]
      \begin{minipage}{.5\linewidth}
            \caption{Training parameters for Bayesian Ridge regressor}\label{table_bayesian_training_params}%
            \centering
                  \begin{tabular}{@{}ll@{}}
                  \toprule
                  Parameter                     & Value \\
                  \midrule
                  Maximum iterations            & 300   \\
                  Tolerance\footnotemark[1]     & \num{1e-03} \\
                  alpha 1                       & \num{1e-06} \\
                  alpha 2                       & \num{1e-06} \\
                  lambda 1                      & \num{1e-06} \\
                  lambda 2                      & \num{1e-06} \\
                  \botrule
                  \end{tabular}
            \footnotetext[1]{Tolerance for the stopping criteria.}
            \caption{Training parameters for GPT-2 regressor}\label{table_gpt2_training_params}%
            \centering
                  \begin{tabular}{@{}ll@{}}
                  \toprule
                  Parameter                     & Value \\
                  \midrule
                  Batch size                    & 10 \\
                  Problem type                  & regression \\
                  Epochs                        & 10 \\
                  Sequence length (tokens)      & 256 \\
                  Tokenizer                     & GPT-2 tokenizer \\
                  \botrule
                  \end{tabular}
      \end{minipage}
      \begin{minipage}{.5\linewidth}
            \caption{Training parameters for XGBoost regressor}\label{table_xgboost_training_params}%
            \centering
                  \begin{tabular}{@{}ll@{}}
                  \toprule
                  Parameter               & Value \\
                  \midrule
                  objective               & reg:squarederror  \\
                  base score              & 0.5 \\
                  booster                 & gbtree  \\
                  colsample bylevel       & 1 \\
                  colsample bynode        & 1 \\
                  colsample bytree        & 1 \\
                  gamma\footnotemark[1]   & 0 \\
                  learning rate           & 0.3 \\
                  max delta step          & 0 \\
                  max depth               & 6 \\
                  min child weight        & 1 \\
                  estimators              & 200  \\
                  n jobs                  & 12  \\
                  num parallel tree       & 1 \\
                  predictor               & auto  \\
                  random state            & 0 \\
                  reg alpha               & 0 \\
                  reg lambda              & 1 \\
                  scale pos weight        & 1 \\
                  subsample               & 2 \\
                  tree method             & auto  \\
                  \botrule
                  \end{tabular}
            \footnotetext[1]{Tolerance for the stopping criteria.}
      \end{minipage}
\end{table}

Similarly, for the boosted tree regressor, we used the default parameters set in the \emph{xgboost} Python library.
We configured the boosted tree regressor model with the training parameters listed in table \ref{table_xgboost_training_params}.

@@

\textcolor{red}{TODO: Create diagram}

The preceding models require specific input formats.
In particular, the boosted tree and bayesian ridge regressors require structured data (e.g a set of predictor features and a set of target variables), whereas GPT-2 expects text strings as input.

@@

% Tags are converted to text tokens. Columns represent token positions, and cells contain the token at a particular position, for a track.
% To tokenize tags, we have used the GPT-2 tokenizer.

@@

Similarly to the tags tabular format, we also defined fixed values for the number of columns: 10, 1,000, and 10,000.
\textcolor{red}{Describe the name of this particular limit (maxsequence length?)}

@@

@@

@@

The experiments tested the three models, and the three values of $K$: \num{100}, \num{1000}, \num{10000}.
In the Bayesian Ridge and Boosted tree regressors, we tested the two structured data formats: tags and tokens as columns.
In the GPT-2 model, we tested the text data format.

@@

\textcolor{red}{TODO: as you have std deviations for your RMSE, can you say that
.291 is really better than .297 ? (idem for other results)}

@@

The GPT-2 model was the best model to predict the danceability, with a \num{0.145} RMSE.
The Boosted tree was the best model to predict the instrumentalness, with a RMSE value of \num{0.29}.

@@

The Duplicated stringification method consistently performed worse with the GPT-2 transformer model.
